---
title: "Class Project"
author: "Sean Morris"
date: "2023-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
setwd("~/Desktop/regression/Class Project")
```


```{r}
# Packages
#install.packages("tidyverse")
#install.packages("readxl")
library(tidyverse)
library(readxl)
library(ggplot2)
library(car)
```

## Data Pre-processing - Ownership Data ##
```{r}
# Gun-ownership db
ownership_init = read_xlsx("data/firearm_ownership.xlsx", sheet = "dataset")

# Dropping unncessecary variables
ownership_init = ownership_init[, -c(1, 5, 8:20)]
ownership_init = ownership_init %>% rename(prop_owners = HFR)

# Calculating National Average for visualization purposes
nat_avg_prop_owners = ownership_init %>% 
                   group_by(Year) %>% 
                   summarize(nat_avg_prop_owners = mean(prop_owners, na.rm = TRUE))

# Now filtering out all states besides CA, NC, & TX
ownership = ownership_init %>% 
                      filter(STATE %in% c("California", "North Carolina", "Texas"))
# Adding Yearly Average
ownership = ownership %>%
   left_join(nat_avg_prop_owners, by = "Year")
```

## Data Pre-processing - Gun Laws Data ##
```{r}
# Gun Laws db
laws_init = read_xlsx("data/firearm_laws.xlsx", sheet = "dataset")
# Dropping unnecessary variables
laws_init = laws_init[, -c(1,3:5,8,11,12,14:27)]
# Now filtering out all states besides CA, NC, & TX
laws_init = laws_init %>%
  filter(State %in% c("California", "North Carolina", "Texas"))
# Filtering by law class, effect (restrictive), and implementation
laws_init = laws_init %>%
  filter(laws_init$`Law Class` %in% c("carrying a concealed weapon (ccw)", "child access laws", "permit to purchase"))

laws_init = laws_init %>%
  filter(laws_init$Effect %in% c("Restrictive"))

laws_init = laws_init %>%
  filter(laws_init$`Type of Change` %in% c("Implement"))
# Filtering by years that fall in other dataframe (1980-2016)
laws <- laws_init %>%
  filter(`Effective Date Year` >= 1980)
```

## Data Pre-processing - Merging Predictors Data ##
```{r}
# Changing Varnames before joining
ownership = ownership %>% rename(State = STATE)
laws = laws %>% rename(Year = `Effective Date Year`)

# Joining together dataframes, while binary-encoding for years before and after the child law is implemented in 'laws'
child_law_years = laws %>%
  filter(`Law Class` == "child access laws") %>%
  select(State, Year)

generate_child_dummy = function(year, state, child_law_years) {
  # Check if state exists in child_law_years
  if (state %in% child_law_years$State) {
    # Get the year of law implementation
    law_year = child_law_years %>%
      filter(State == state) %>%
      pull(Year)
    # Compare the year from 'ownership' with the year from 'law'
    return(ifelse(year >= law_year, 1, 0))
  } else {
    # If state not found, return 0
    return(0)
  }
}

# Applying function to each row in "ownership"
ownership$child = mapply(generate_child_dummy,
                         ownership$Year,
                         ownership$State,
                         MoreArgs = list(child_law_years = child_law_years))

# Creating final regressor df, "guns" 
guns = ownership %>%
  select(Year, State, prop_owners, nat_avg_prop_owners, everything())
```


## Data Pre-processing - Gun violence database ##
```{r}
v_1980_to_1998 = read.table("data/v_1980-1998.txt", header = TRUE, sep = "\t", quote = "", fill = TRUE)
v_1999_to_2016 = read.table("data/v_1999-2016.txt", header = TRUE, sep = "\t", quote = "", fill = TRUE)

# Dropping "notes" Column
v_1980_to_1998 <- v_1980_to_1998 %>% select(-X.Notes.)
v_1999_to_2016 <- v_1999_to_2016 %>% select(-X.Notes.)

# Dropping NA rows
v_1980_to_1998 <- v_1980_to_1998 %>% 
  filter(!is.na(X.State.) & X.State. != "")
v_1999_to_2016 <- v_1999_to_2016 %>% 
  filter(!is.na(X.State.) & X.State. != "")

# Making sure there are no more null entries
cat(any(is.na(v_1980_to_1998)),any(is.na(v_1999_to_2016)))

# Dropping unecessary variables 
v_1980_to_1998 = v_1980_to_1998[, -c(2,4:6)]
v_1999_to_2016 = v_1999_to_2016[, -c(2,4:6)]


# Checking Unique Age Identifiers before aggregating
unique(v_1980_to_1998$X.Age.Group.)
unique(v_1999_to_2016$X.Age.Group.)
  # Looks like there is an entry called "Not Stated" in 1980-1998 data - lets drop it 
v_1980_to_1998 <- v_1980_to_1998 %>%
  filter(X.Age.Group. != "\"Not Stated\"")
  # Checking again
unique(v_1980_to_1998$X.Age.Group.)


# Reconstructing Dataframes to get single-year rows for modelling
through_1998 = v_1980_to_1998
through_2016 = v_1999_to_2016
  # Categorizing minors and adults
through_1998 <- through_1998 %>%
  mutate(Age_Category = case_when(
    str_trim(X.Age.Group.) %in% c("\"< 1 year\"", "\"1-4 years\"", "\"5-9 years\"", "\"10-14 years\"", "\"15-19 years\"") ~ "minors",
    TRUE ~ "adults"
  ))
through_2016 <- through_2016 %>%
  mutate(Age_Category = case_when(
    str_trim(X.Age.Group.) %in% c("\"< 1 year\"", "\"1-4 years\"", "\"5-9 years\"", "\"10-14 years\"", "\"15-19 years\"") ~ "minors",
    TRUE ~ "adults"
  ))

# Converting 'Population' column to numeric, handling non-numeric entries
through_1998$Population <- as.numeric(gsub("\"", "", through_1998$Population))
through_2016$Population <- as.numeric(gsub("\"", "", through_2016$Population))

# Also convert the 'Deaths' column to numeric if it's not already
through_1998$Deaths <- as.numeric(gsub("\"", "", through_1998$Deaths))
through_2016$Deaths <- as.numeric(gsub("\"", "", through_2016$Deaths))

# Creating Population and Death dfs by state, year, and age category in order to calculate rates later
pop_1998 = through_1998 %>%
  select(X.State., X.Year., Population, Age_Category) %>%
  distinct() %>%
  group_by(X.State., X.Year., Age_Category) %>%
  summarize(
    total_population = sum(Population, na.rm = TRUE),
    .groups = 'drop'
  )

pop_2016 = through_2016 %>%
  select(X.State., X.Year., Population, Age_Category) %>%
  distinct() %>%
  group_by(X.State., X.Year., Age_Category) %>%
  summarize(
    total_population = sum(Population, na.rm = TRUE),
    .groups = 'drop'
  )

pop_1998_wide = pop_1998 %>%
  pivot_wider(
    id_cols = c(X.State., X.Year.),
    names_from = Age_Category,
    values_from = total_population,
    names_prefix = "total_",
    names_sep = "_"
  )

pop_2016_wide = pop_2016 %>%
  pivot_wider(
    id_cols = c(X.State., X.Year.),
    names_from = Age_Category,
    values_from = total_population,
    names_prefix = "total_",
    names_sep = "_"
  )

death_1998 = through_1998 %>%
  group_by(X.State., X.Year., X.Injury.Intent., Age_Category) %>%
  summarize(
    deaths = sum(Deaths, na.rm = TRUE),
    .groups = 'drop'
  )

death_2016 = through_2016 %>%
  group_by(X.State., X.Year., X.Injury.Intent., Age_Category) %>%
  summarize(
    deaths = sum(Deaths, na.rm = TRUE),
    .groups = 'drop'
  )

through_1998_aggregate = left_join(death_1998, pop_1998, by = c("X.State.", "X.Year.", "Age_Category"))
through_2016_aggregate = left_join(death_2016, pop_2016, by = c("X.State.", "X.Year.", "Age_Category"))

through_1998_aggregate = through_1998_aggregate %>%
  mutate(
    rate_per_100k = (deaths / total_population) * 100000
  )

through_2016_aggregate = through_2016_aggregate %>%
  mutate(
    rate_per_100k = (deaths / total_population) * 100000
  )

# Creating aggregate DF
through_1998_aggregate = through_1998 %>%
  group_by(X.State., X.Year., X.Injury.Intent., Age_Category) %>%
  summarize(
    deaths = sum(Deaths, na.rm = TRUE),
    population = sum(Population, na.rm = TRUE),
    rate_per_100k = (sum(Deaths, na.rm = TRUE) / sum(Population, na.rm = TRUE)) * 100000,
    .groups = 'drop'
  )

through_2016_aggregate = through_2016 %>%
  group_by(X.State., X.Year., X.Injury.Intent., Age_Category) %>%
  summarize(
    deaths = sum(Deaths, na.rm = TRUE),
    population = sum(Population, na.rm = TRUE),
    rate_per_100k = (sum(Deaths, na.rm = TRUE) / sum(Population, na.rm = TRUE)) * 100000,
    .groups = 'drop'
  )



# Pivoting
through_1998_aggregate_wide = through_1998_aggregate %>%
  pivot_wider(
    id_cols = c(X.State., X.Year.),
    names_from = c(X.Injury.Intent., Age_Category),
    values_from = c(deaths, rate_per_100k),
    names_sep = "_"
  )

through_2016_aggregate_wide = through_2016_aggregate %>%
  pivot_wider(
    id_cols = c(X.State., X.Year.),
    names_from = c(X.Injury.Intent., Age_Category),
    values_from = c(deaths, rate_per_100k),
    names_sep = "_"
  )

  # replacing NA values with 0, mostly for unintended deaths, which did not occur in certain years
through_1998_aggregate_wide[] <- lapply(through_1998_aggregate_wide, function(x) ifelse(is.na(x), 0, x))
through_2016_aggregate_wide[] <- lapply(through_2016_aggregate_wide, function(x) ifelse(is.na(x), 0, x))


# Some housekeeping with variable names and dtypes
through_1998_aggregate_wide = through_1998_aggregate_wide %>%
  rename(State = X.State., Year = X.Year.) %>%
  mutate(
    State = gsub("\"", "", State),
    Year = as.numeric(gsub("\"", "", Year))
  )

through_2016_aggregate_wide = through_2016_aggregate_wide %>%
  rename(State = X.State., Year = X.Year.) %>%
  mutate(
    State = gsub("\"", "", State),
    Year = as.numeric(gsub("\"", "", Year))
  )

# Combining violence Data
violence_init = bind_rows(through_1998_aggregate_wide, through_2016_aggregate_wide)
violence = violence_init %>%
  arrange(State, Year)

# Dealing with Null values again
violence[] <- lapply(violence, function(x) ifelse(is.na(x), 0, x))


#### Adding back population variables ####
# First fixing varnames and dtypes
pop_1998_wide = pop_1998_wide %>%
  rename(State = X.State., Year = X.Year.) %>%
  mutate(
    State = gsub("\"", "", State),
    Year = as.numeric(gsub("\"", "", Year))
  )

pop_2016_wide = pop_2016_wide %>%
  rename(State = X.State., Year = X.Year.) %>%
  mutate(
    State = gsub("\"", "", State),
    Year = as.numeric(gsub("\"", "", Year))
  )

# Adding total population variable to each 
pop_1998_wide <- pop_1998_wide %>%
  mutate(total_pop = total_adults + total_minors)

pop_2016_wide <- pop_2016_wide %>%
  mutate(total_pop = total_adults + total_minors)

# Combining
population = bind_rows(pop_1998_wide, pop_2016_wide)
population = population %>%
  arrange(State, Year)
```


## Data Pre-processing - Baseline Dataframe ##
```{r}
data_init = left_join(guns, violence, by = c("State", "Year"))
data = data_init %>%
  left_join(population, by = c("State", "Year"))

data <- data %>%
  rename(
    adult_homicides = `deaths_"Homicide"_adults`,
    minor_homicides = `deaths_"Homicide"_minors`,
    adult_suicides = `deaths_"Suicide"_adults`,
    minor_suicides = `deaths_"Suicide"_minors`,
    adult_unintentionals = `deaths_"Unintentional "_adults`,
    minor_unintentionals = `deaths_"Unintentional "_minors`,
    adult_homicide_rate = `rate_per_100k_"Homicide"_adults`,
    minor_homicide_rate = `rate_per_100k_"Homicide"_minors`,
    adult_suicide_rate = `rate_per_100k_"Suicide"_adults`,
    minor_suicide_rate = `rate_per_100k_"Suicide"_minors`,
    adult_unintentional_rate = `rate_per_100k_"Unintentional "_adults`,
    minor_unintentional_rate = `rate_per_100k_"Unintentional "_minors`
  )

data = data[, -c(20:23)]

data <- data %>%
  mutate(
    total_homicides = adult_homicides + minor_homicides,
    total_suicides = adult_suicides + minor_suicides,
    total_unintentionals = adult_unintentionals + minor_unintentionals,
    total_homicide_rate = (total_homicides / total_pop) * 100000,
    total_suicide_rate = (total_suicides / total_pop) * 100000,
    total_unintentional_rate = (total_unintentionals / total_pop) * 100000)

```

## Data Pre-processing - Additional Variables ##
```{r}
employment = read_xls('data/UNRATE_1980_2016.xls', sheet = 'dataset')
data = left_join(data, employment, by = c("Year"))
single_parents = read_xls('data/single_parent_1980_2016.xls', sheet = 'dataset')

data = left_join(data, single_parents, by = c("Year"))

#data_trunc = data %>%
#  filter(Year >= 1999)
#
#full_data = left_join(data_trunc, employment, by = c("State", "Year"))
```


## EDA - Visualization 1 ##
```{r}
p = ggplot(data, aes(x=Year, y = prop_owners, color = State)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Proportion of Gun Owners By State",
       x = "Year",
       y = "% of Population Owning a Firearm") +
  scale_color_brewer(palette = "Set1")

p + geom_line(data=data %>% distinct(Year, nat_avg_prop_owners),
              aes(x = Year, y = nat_avg_prop_owners, color = "National Average"),
              linetype = "dashed")
```

## EDA - Visualization 2 ##

```{r}
ratio <- max(data$total_homicide_rate) / max(data$prop_owners)

ggplot(data, aes(x = Year)) +
  geom_line(aes(y = total_homicide_rate, color = "Homicide Rate"), size = 1) +
  geom_line(aes(y = prop_owners * ratio, color = "% of Gun Owners"), size = 1) +
  facet_wrap(~ State) +
  labs(title = "Total Homicide Rate and \nProportion of Gun Owners Over Time by State",
       x = "Year",
       y = "Total Homicide Rate") +
  scale_y_continuous(
    # Primary axis
    name = "Total Homicide Rate",
    sec.axis = sec_axis(~ . / ratio, name = "Proportion of Gun Owners")  # Secondary axis
  ) +
  scale_color_manual(values = c("Homicide Rate" = "red", "% of Gun Owners" = "blue")) +
  theme_minimal() +
  theme(text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(color = "red"), 
        axis.title.y.right = element_text(color = "blue")  
  )

```

## EDA - Visualization 3 ##

```{r}
ratio <- max(data$total_suicide_rate) / max(data$prop_owners)

ggplot(data, aes(x = Year)) +
  geom_line(aes(y = total_suicide_rate, color = "Suicide Rate"), size = 1) +
  geom_line(aes(y = prop_owners * ratio, color = "% of Gun Owners"), size = 1) +
  facet_wrap(~ State) +
  labs(title = "Total Suicide Rate and \nProportion of Gun Owners Over Time by State",
       x = "Year",
       y = "Total Suicide Rate",
       fill = "Metric") +
  scale_y_continuous(
    # Primary axis
    name = "Total Suicide Rate",
    sec.axis = sec_axis(~ . / ratio, name = "Proportion of Gun Owners")  # Secondary axis
  ) +
  scale_color_manual(values = c("Suicide Rate" = "orange", "% of Gun Owners" = "blue")) +
  theme_minimal() +
  theme(text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        axis.title.y = element_text(color = "orange"), 
        axis.title.y.right = element_text(color = "blue")  
  )
```


## EDA - Visualization 4 ##
```{r}
ggplot(data, aes(x = Year)) +
  geom_col(aes(y = adult_suicides, fill = "adult_suicides"), position = "stack") +
  geom_col(aes(y = minor_suicides, fill = "minor_suicides"), position = "stack") +
  facet_wrap(~ State) +
  labs(title = "Suicides by Age Group Over Time by State",
       x = "Year",
       y = "Number of Suicides",
       fill = "Age Group") +
  scale_fill_manual(values = c("adult_suicides" = "blue", "minor_suicides" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))  # Rotate x-axis labels for readability

```

## EDA - Visualization 5 ##
```{r}
ggplot(data, aes(x = Year)) +
  geom_col(aes(y = adult_homicides, fill = "adult_homicides"), position = "stack") +
  geom_col(aes(y = minor_homicides, fill = "minor_homicides"), position = "stack") +
  facet_wrap(~ State) +
  labs(title = "Homicides by Age Group Over Time by State",
       x = "Year",
       y = "Number of Homicides",
       fill = "Age Group") +
  scale_fill_manual(values = c("adult_homicides" = "blue", "minor_homicides" = "red")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))  # Same as above

```

## EDA - Visaulization 6 ##
```{r}
# Create the plot grid with homicide_rate and add pct_change_employment as points
ggplot(data, aes(x = Year)) +
  geom_line(aes(y = total_homicide_rate, group = State, color = "Homicide Rate"), size = 1) +
  geom_line(aes(y = urate, color = "Unemployment Rate"), size = 0.5) +
  facet_wrap(~ State, scales = 'free_y') +
  labs(title = "Homicide Rates and National Unemployment",
       x = "Year",
       y = "Metric",
       color = "Metric") +
  scale_color_manual(values = c("Homicide Rate" = "red", "Unemployment Rate" = "black")) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

## EDA - Visualization 7 ##
```{r}
# Create the plot grid with homicide_rate and add pct_change_employment as points
ggplot(data, aes(x = Year)) +
  geom_line(aes(y = total_suicide_rate, group = State, color = "Suicide Rate"), size = 1) +
  geom_line(aes(y = urate, color = "Unemployment Rate"), size = 0.5) +
  facet_wrap(~ State, scales = 'free_y') +
  labs(title = "Suicide Rates and National Unemployment",
       x = "Year",
       y = "Metric",
       color = "Metric") +
  scale_color_manual(values = c("Suicide Rate" = "orange", "Unemployment Rate" = "black")) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## EDA - Looking at each variable's log-transform
```{r}
for(state in c("California", "North Carolina", "Texas")) {
  for(variable in c("prop_owners", "total_homicide_rate", "total_suicide_rate", "total_unintentional_rate")) {
    state_data <- data[data$State == state, ]
    if (nrow(state_data) > 0) {
      state_data$log_variable <- log(state_data[[variable]])  

      p <- ggplot(state_data, aes_string(x = "Year")) +
        geom_line(aes_string(y = variable, colour = "'Original'")) +
        geom_line(aes(y = log_variable, colour = "'Log-Transformed'")) +
        labs(title = paste("Variable:", variable, "in", state), y = "Value", colour = "Legend") +
        theme_minimal()
      print(p)
    }
  }
}



```
```{r}
for(variable in c("urate", "total_single_parent")) {
  data$log_variable <- log(data[[variable]])  

  p <- ggplot(data, aes_string(x = "Year")) +
    geom_line(aes_string(y = variable, colour = "'Original'")) +
    geom_line(aes(y = log_variable, colour = "'Log-Transformed'")) +
    labs(title = paste("Variable:", variable), y = "Value", colour = "Legend") +
    theme_minimal()

  print(p) 
}
```

From this info above - I will elect to use log-transforms of the respective homicide/suicide/unintentional rates, as well as the log of unemployment, while keeping single_parent children and prop_owners the same. 

## Pre-Modelling Tests
```{r}
# Testing for non-stationarity 
#install.packages("tseries")
#install.packages("lmtest")
#install.packages("sandwich")
#install.packages("plm")
library(tseries)
library(lmtest)
library(sandwich)
library(plm)

adf.test(log(data$total_homicide_rate)) #Non-stationary
adf.test(log(data$total_suicide_rate)) #Non-stationary
adf.test(data$prop_owners) #Non-stationary
adf.test(log(data$urate)) #Stationary
adf.test(data$total_single_parent) #Non-stationary
```


## Model A - Testing Policy impact of permit to purchase on total suicide rates by state 
```{r}
model1_df <- data %>%
  select(Year, State, prop_owners, permit, adult_suicide_rate, minor_suicide_rate,
         total_suicide_rate, urate, total_single_parent, single_mother, single_father)

model_0 = lm(total_suicide_rate ~ permit + urate + total_single_parent, data = model1_df)
print((summary(model_0)))
# Problem - Our variables are non-stationary, and since we are working with panel data it would be wise to use clustered standard errors as well. Finally, lets log-transform our variables to better smooth them out. 

#Log-transforming and differencing variables
model1_df <- model1_df %>%
  mutate(log_total_suicide_rate = log(total_suicide_rate),
         log_prop_owners = log(prop_owners),
         log_urate = log(urate),
         log_total_single_parent = log(total_single_parent),
         diff_log_total_suicide_rate = c(NA, diff(log_total_suicide_rate)),
         diff_log_prop_owners = c(NA, diff(log_prop_owners)),
         diff_log_urate = c(NA, diff(log_urate)),
         diff_log_total_single_parent = c(NA, diff(log_total_single_parent)))

# Running a log-log model with no differences
model_1 = lm(log_total_suicide_rate ~
               permit + 
               log_prop_owners + 
               log_urate + 
               log_total_single_parent, 
               data = model1_df)
print((summary(model_1)))

    # Looking better, but I will now add clustered standard errors to account for any group-dependence
cluster_se <- coeftest(model_1, vcov = vcovHC(model_1, type = "HC1", cluster = "group", group = model1_df$State))
print(cluster_se)

# ok, now lets run the same regression using first differences of each non-stationary variable
model_2 = lm(diff_log_total_suicide_rate ~ 
               permit +
               diff_log_prop_owners +
               log_urate +
               diff_log_total_single_parent,
             data = model1_df)
print((summary(model_2)))

  # Now with clustered SEs
cluster_se2 <- coeftest(model_2, vcov = vcovHC(model_2, type = "HC1", cluster = "group", group = model1_df$State))
print(cluster_se2)



# Looks like model 1 performs the best.
```
```{r}

```


## Model A - Visualized 
```{r}
# Calculate fitted values
model1_df$fitted_values = fitted(model_1)

# Plotting
ggplot(model1_df, aes(x = log(total_suicide_rate), y = fitted_values)) +
  geom_point(aes(color = State), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ State) +  # Separate plots for each state
  labs(title = "Actual vs Fitted Log(Total Suicide Rate) by State",
       x = "Log of Actual Total Suicide Rate",
       y = "Fitted Values",
       color = "State") +
  theme_minimal()
```

## Model A - Testing for multicollinearity 
```{r}
vif(model_1)
```

## Model A - Looking for potential heteroskedasticity
```{r}
model1_df$std_residuals = rstandard(model_1)

ggplot(model1_df, aes(x = log(prop_owners), y = std_residuals)) +
  geom_point(aes(color = State), alpha = 0.6) +
  geom_hline(yintercept = 0, linetype='dashed', color='red') +
  facet_wrap(~ State) +
  labs(title = "Standardized Residuals vs Log(Prop Owners) by State",
       x = "Log(Prop Owners)",
       y = "Standardized Residuals",
       color = "State") +
  theme_minimal()
```
```{r}
ggplot(model1_df, aes(x = log(urate), y = std_residuals)) +
  geom_point(aes(), alpha = 0.6) +
  geom_hline(yintercept = 0, linetype='dashed', color='red') +
  labs(title = "Standardized Residuals vs Log(urate)",
       x = "Log of Unemployment Rate",
       y = "Standardized Residuals") +
  theme_minimal()
```

```{r}
ggplot(model1_df, aes(x = log(total_single_parent), y = std_residuals)) +
  geom_point(aes(), alpha = 0.6) +
  geom_hline(yintercept = 0, linetype='dashed', color='red') +
  labs(title = "Standardized Residuals vs Log(total_single_parent)",
       x = "Log of # Children in Single-parent Households",
       y = "Standardized Residuals") +
  theme_minimal()
```

## Model A - Aside (looking at variables' impact in minor_suicide_rate)

```{r}
# Running a log-log model with no differences
model_minors = lm(log(minor_suicide_rate) ~
               permit + 
               log_prop_owners + 
               log_urate + 
               log_total_single_parent, 
               data = model1_df)
print((summary(model_minors)))

    # Looking better, but I will now add clustered standard errors to account for any group-dependence
cluster_se <- coeftest(model_1, vcov = vcovHC(model_1, type = "HC1", cluster = "group", group = model1_df$State))
print(cluster_se)

# Pretty Similar
```





## Model B - Testing Policy impact of universal background checks on total homicide rates by state
```{r}
model2_df <- data %>%
  select(Year, State, prop_owners, universl, adult_homicide_rate, minor_homicide_rate,
         total_homicide_rate, urate, total_single_parent, single_mother, single_father)

model2_df <- model2_df %>%
  mutate(log_total_homicide_rate = log(total_homicide_rate),
         log_prop_owners = log(prop_owners),
         log_urate = log(urate),
         log_total_single_parent = log(total_single_parent))


# Running a log-log model with no differences
model_3 = lm(log_total_homicide_rate ~
               universl + 
               log_prop_owners + 
               log_urate + 
               log_total_single_parent, 
               data = model2_df)
print((summary(model_3)))

    # Looking better, but I will now add clustered standard errors to account for any group-dependence
cluster_se3 <- coeftest(model_3, vcov = vcovHC(model_3, type = "HC1", cluster = "group", group = model2_df$State))
print(cluster_se3)
```


## Model B - Visualized
```{r}
# Calculate fitted values
model2_df$fitted_values = fitted(model_3)

# Plotting
ggplot(model2_df, aes(x = log_total_homicide_rate, y = fitted_values)) +
  geom_point(aes(color = State), alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ State) +  # Separate plots for each state
  labs(title = "Actual vs Fitted Log(Total Homicide Rate) by State",
       x = "Log of Total Homicide Rate",
       y = "Fitted Values",
       color = "State") +
  theme_minimal()
```

## Model B - Testing for multicollinearity 
```{r}
vif(model_3)
```

## Model B - Looking for potential heteroskedasticity
```{r}
model2_df$std_residuals = rstandard(model_3)

ggplot(model2_df, aes(x = log(prop_owners), y = std_residuals)) +
  geom_point(aes(color = State), alpha = 0.6) +
  geom_hline(yintercept = 0, linetype='dashed', color='red') +
  facet_wrap(~ State) +
  labs(title = "Standardized Residuals vs Log(Prop Owners) by State",
       x = "Log(Prop Owners)",
       y = "Standardized Residuals",
       color = "State") +
  theme_minimal()
```
```{r}
ggplot(model2_df, aes(x = log(urate), y = std_residuals)) +
  geom_point(aes(), alpha = 0.6) +
  geom_hline(yintercept = 0, linetype='dashed', color='red') +
  labs(title = "Standardized Residuals vs Log(urate)",
       x = "Log of Unemployment Rate",
       y = "Standardized Residuals") +
  theme_minimal()
```
```{r}
ggplot(model2_df, aes(x = log(total_single_parent), y = std_residuals)) +
  geom_point(aes(), alpha = 0.6) +
  geom_hline(yintercept = 0, linetype='dashed', color='red') +
  labs(title = "Standardized Residuals vs Log(total_single_parent)",
       x = "Log of # Children in Single-parent Households",
       y = "Standardized Residuals") +
  theme_minimal()
```



